---
title: "Buffering PAD"
author: "Elmera Azadpour"
date: "5/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(fasterize)
library(raster)
library(tidyverse)
library(rgdal)
library(here)
library(terra)
```

## read in PAD data
```{r}
PAD_GAP_1_2_eco4<- sf::st_read(here("PAD_dissolve_clip","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US_dissolve_Unit_Nm_clip_eco4.shp")) %>% 
  dplyr::select(GAP_Sts) 

PAD_df_filt_96km<- sf::st_read(here("PAD_df_filt","PAD_df_BCrange_96km.shp")) %>% 
  dplyr::select(GAP_Sts) 
```

## set crs so when we buffer we know the units (meters)
```{r}
st_crs(PAD_GAP_1_2_eco4) <-  "EPSG:5070"
st_crs(PAD_GAP_1_2_eco4)

st_crs(PAD_df_filt_96km) <-  "EPSG:5070"
st_crs(PAD_df_filt_96km)
```

# buffer the entire PAD GAP 1&2 SE ecoregion polygons by 10km and then buffer the 96km bobcat range shapefile
```{r}
# PAD_GAP_1_2_eco4_10km <- st_buffer(PAD_GAP_1_2_eco4, 10000)

# PAD_df_filt_96km_10km <- st_buffer(PAD_df_filt_96km, 10000)


# add buffer, fortunately the base CRS is in meters
PAD_df_filt_96km_buffered <- st_buffer(PAD_df_filt_96km, 10000) 

# polygon(s) of only the buffer
buffer_only <- st_difference(PAD_df_filt_96km_buffered, st_union(PAD_df_filt_96km))

plot(buffer_only[1,1])
plot(PAD_df_filt_96km_buffered[1,1])

# sf::st_write(buffer_only,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_only_10km/buffer_only_10km.shp") # export df as shp to view in qgis 
```


## now lets apply the FORESCE 2050 A1B projection with the 4 classes to the buffered vector
## our goal here is to see land use change pressure nearby these core areas 
## see if can calucate % of the different classes in these buffer regions
```{r}
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_filt_96km_10km_buff <- sf::st_read(here("PAD_df_filt_buffer_10km","PAD_df_BCrange_96km_10km.shp"))

PAD_df_filt_96km_10km_buff <- st_transform(
  PAD_df_filt_96km_10km_buff,
  crs=st_crs(template)
)

PAD_df_filt_96km_10km_buff$GAP_Sts = as.numeric(PAD_df_filt_96km_10km_buff$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_filt_96km_10km_buff,
  raster = template,
field = "GAP_Sts")

 plot(raster)

# raster::writeRaster(
#    raster,
#    "PAD_df_filt_96km_10km_buff.tif",
#    overwrite = TRUE
#    )

## FORESCE LULC
PAD_filt_96km_10km_buff <- rast(here("PAD_df_filt_buffer_10km", "PAD_df_filt_96km_10km_buff.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt_96km_10km_buff) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt_96km_10km_buff) %>%
  mask(mask = PAD_filt_96km_10km_buff)

plot(foresce_pad_filt)

# save raster as .tif
# terra::writeRaster(
#   foresce_pad_filt,
#   "FORESCE_2050_A1B_us_extent_96km_buff_10km.tif",
#   overwrite = TRUE
#   )

## now crop to SE US
FORESCE_2050_A1B_us_extent_96km_buff_10km <- rast(here("PAD_df_filt_buffer_10km", "FORESCE_2050_A1B_us_extent_96km_buff_10km.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_96km_buff_10km) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_96km_buff_10km %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

 plot(crop)

# save raster as .tif
# terra::writeRaster(
#   crop,
#   "FORESCE_2050_A1B_eco4_l3_96km_buff_10km.tif",
#   overwrite = TRUE
#   )
```

## cell stats
```{r}
FORESCE_2050_A1B_eco4_l3_96km_buff_10km <- rast(here("PAD_df_filt_buffer_10km", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km.tif"))


hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km, 
        ylim=c(0,15000),
     xaxp = c(0.0, 3.0, 1.0),
     breaks = 5)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km,
        ylim=c(0,45000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer")


FORESCE_2050_A1B_eco4_l3_96km <- rast(here("PAD_df_filt_BC_96km", "FORESCE_2050_A1B_eco4_l3_96km_filt.tif"))

barplot(FORESCE_2050_A1B_eco4_l3_96km,
        ylim=c(0,8000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km")



```

## look at certain larger polygons and what the buffer region around them consists of in terms of % natural, cropland, etc 
## you'll need to filter 
```{r}
## read in BC range 96km shp and filter for larger polygons then buffer and look at the pixels in those buffer regions 
PAD_df_filt_96km <- sf::st_read(here("PAD_df_filt","PAD_df_BCrange_96km.shp"))

PAD_df_96km_Apalachicola <- PAD_df_filt_96km %>% 
  filter(Unit_Nm %in% c("Apalachicola National Estuarine Research Reserve"))

PAD_df_96km_Rookery <- PAD_df_filt_96km %>% 
  filter(Unit_Nm %in% c("Rookery Bay National Estuarine Research Reserve"))

PAD_df_96km_Land_Between <- PAD_df_filt_96km %>% 
  filter(Unit_Nm %in% c("Land Between The Lakes Wildlife Management Area"))

PAD_df_96km_Cohutta <- PAD_df_filt_96km %>% 
  filter(Unit_Nm %in% c("Cohutta Wildlife Management Area"))

PAD_df_96km_Kissimmee <- PAD_df_filt_96km %>% 
  filter(Unit_Nm %in% c("Kissimmee Prairie Preserve State Park"))

## now lets buffer these indiviual polygons 
PAD_df_96km_Apalachicola_buff <- st_buffer(PAD_df_96km_Apalachicola, 10000)
PAD_df_96km_Rookery_buff <- st_buffer(PAD_df_96km_Rookery, 10000)
PAD_df_96km_Land_Between_buff <- st_buffer(PAD_df_96km_Land_Between, 10000)
PAD_df_96km_Cohutta_buff <- st_buffer(PAD_df_96km_Cohutta, 10000)
PAD_df_96km_Kissimmee_buff <- st_buffer(PAD_df_96km_Kissimmee, 10000)

## now lets export each of these indiual buffered polygons 
# sf::st_write(PAD_df_96km_Apalachicola_buff,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_top_polygons/PAD_df_BCrange_96km_10km_Apalachicola.shp") 
# 
# sf::st_write(PAD_df_96km_Rookery_buff,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_top_polygons/PAD_df_BCrange_96km_10km_Rookery.shp") 
# 
# sf::st_write(PAD_df_96km_Land_Between_buff,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_top_polygons/PAD_df_BCrange_96km_10km_Land_Between.shp") 
# 
# sf::st_write(PAD_df_96km_Cohutta_buff,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_top_polygons/PAD_df_BCrange_96km_10km_Cohutta.shp") 
# 
# sf::st_write(PAD_df_96km_Kissimmee_buff,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/buffer_top_polygons/PAD_df_BCrange_96km_10km_Kissimmee.shp") 

## now lets rasterize and crop and mask to the FORESCE data 
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_96km_Kissimmee_buff <- sf::st_read(here("buffer_top_polygons","PAD_df_BCrange_96km_10km_Kissimmee.shp"))

PAD_df_96km_Kissimmee_buff <- st_transform(
  PAD_df_96km_Kissimmee_buff,
  crs=st_crs(template)
)

PAD_df_96km_Kissimmee_buff$GAP_Sts = as.numeric(PAD_df_96km_Kissimmee_buff$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_96km_Kissimmee_buff,
  raster = template,
field = "GAP_Sts")

# raster::writeRaster(
#    raster,
#    "PAD_df_filt_96km_10km_buff_Kissimmee.tif",
#    overwrite = TRUE
#    )

###################################### 
############  FORESCE LULC ############  
###################################### 

PAD_filt_96km_10km_buff_Kissimmee<- rast(here("buffer_top_polygons_raster", "PAD_df_filt_96km_10km_buff_Kissimmee.tif"))

FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt_96km_10km_buff_Kissimmee) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt_96km_10km_buff_Kissimmee) %>%
  mask(mask = PAD_filt_96km_10km_buff_Kissimmee)

# plot(foresce_pad_filt)

# save raster as .tif
# terra::writeRaster(
#   foresce_pad_filt,
#   "FORESCE_2050_A1B_us_extent_96km_buff_10km_Kissimmee.tif",
#   overwrite = TRUE
#   )


## next step 
## now crop to SE US
FORESCE_2050_A1B_us_extent_96km_buff_10km_Kissimmee <- rast(here("buffer_top_polygons_raster", "FORESCE_2050_A1B_us_extent_96km_buff_10km_Kissimmee.tif"))

FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_96km_buff_10km_Kissimmee) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_96km_buff_10km_Kissimmee %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

# plot(crop)

# save raster as .tif
# terra::writeRaster(
#   crop,
#   "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Kissimmee.tif",
#   overwrite = TRUE
#   )

```

## count of how many FORESCE pixels in each of these polygon regions
```{r}
FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola <- rast(here("buffer_top_polygons_raster", "PAD_sites", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola.tif"))

summary(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)


# hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola,
        ylim=c(0,2000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer", 
        sub = "Unit_Nm = Apalachicola National Estuarine Research Reserve, FL" )


FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Cohutta<- rast(here("buffer_top_polygons_raster", "PAD_sites", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Cohutta.tif"))

summary(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Cohutta)

# hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Cohutta,
        ylim=c(0,2000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer", 
        sub = "Unit_Nm =  Cohutta Wildlife Managment Area, GA" )




FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Rookery<- rast(here("buffer_top_polygons_raster", "PAD_sites", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Rookery.tif"))

summary(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Cohutta)

# hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Rookery,
        ylim=c(0,2000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer", 
        sub = "Unit_Nm = Rookery Bay National Estuarine Research Reserve, FL" )

FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Land_Between <- rast(here("buffer_top_polygons_raster", "PAD_sites", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Land_Between.tif"))

summary(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Land_Between)

# hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Land_Between,
        ylim=c(0,2000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer", 
        sub = "Unit_Nm = Land Between the Lakes Wildlife Management Area, TN" )


FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Kissimmee<- rast(here("buffer_top_polygons_raster", "PAD_sites", "FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Kissimmee.tif"))

summary(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Kissimmee)

# hist(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Apalachicola)

barplot(FORESCE_2050_A1B_eco4_l3_96km_buff_10km_Kissimmee,
        ylim=c(0,2000),
        col = c("#ff1212", "#ffff00", "#ed4cdb", "#33a02c"),
        names=c("Other", "Cropland", "Hay/Pasture", "Natural"),
        main = "Bobcat homerange filtered >= 96km with 10km buffer", 
        sub = "Unit_Nm = Kissimmee Prairie Preserve State Park, FL " )


```

