---
title: "reclassification/rasterization"
author: "Elmera Azadpour"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(fasterize)
library(raster)
library(tidyverse)
library(rgdal)
library(here)
library(terra)
```

## st_transform() gap to match foresce raster crs
## fasteruze gap data to final raster for foresce
```{r}
PAD_GAP_1_2_US <- sf::st_read(here("PAD","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US.shp")) %>% 
  dplyr::select(GAP_Sts) 


template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

PAD_GAP_1_2_US <- st_transform(
  PAD_GAP_1_2_US,
  crs=st_crs(template)
)


PAD_GAP_1_2_US$GAP_Sts = as.numeric(PAD_GAP_1_2_US$GAP_Sts)

raster <- fasterize(
  sf = PAD_GAP_1_2_US,
  raster = template,
field = "GAP_Sts")

plot(raster)

# raster::writeRaster(
#    raster,
#    "PAD_GAP_1_2.tif",
#    overwrite = TRUE
#    )
```


## now lets crop and mask the new PAD_GAP_1_2.tif to the FORESCEA1B 2050 eco4 level3 
```{r}
GAP12 <- rast(here("PAD_1_2_rasters", "PAD_GAP_1_2.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

plot(GAP12)
plot(FORESCE_A1B)
# set crs so they're the same 
crs(GAP12) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad <- FORESCE_A1B %>%
  crop(y = GAP12) %>%
  mask(mask = GAP12)

plot(foresce_pad)

# save raster as .tif
terra::writeRaster(
  foresce_pad,
  "FORESCE_2050_A1B_us_extent.tif",
  overwrite = TRUE
  )
```


## now we can crop that new FORESCE_2050_A1B_us_extent.tif to the FORESCE2050A1B_eco_4_Level_3_2013.tif 
```{r}
FORESCE_A1B_eco4lev3_us <- rast(here("FORESCE_2050_A1B_us_extent.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_A1B_eco4lev3_us) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_A1B_eco4lev3_us %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3.tif",
  overwrite = TRUE
  )

```



## Read in filtered area > 10000m2 or (1 ha), > 25000m2 (2.5ha) and > 50000m2 (5ha) - note: filtering was done in qgis
```{r}
PAD_GAP_1_2_US_6.25ha <- sf::st_read(here("PAD_6.25ha_v2","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US_6.25ha_v2.shp")) %>% 
  dplyr::select(GAP_Sts) 

PAD_GAP_1_2_US_6.25ha <- st_transform(
  PAD_GAP_1_2_US_6.25ha,
  crs=st_crs(template)
)


PAD_GAP_1_2_US_6.25ha$GAP_Sts = as.numeric(PAD_GAP_1_2_US_6.25ha$GAP_Sts)

raster_6.5ha <- fasterize(
  sf = PAD_GAP_1_2_US_6.25ha,
  raster = template,
field = "GAP_Sts")

plot(raster_6.5ha)

raster::writeRaster(
   raster_6.5ha,
   "PAD_GAP_1_2_6.25ha_v2.tif",
   overwrite = TRUE
   )
```


# cropping  mask the new PAD_GAP_1_2_1ha.tif,  PAD_GAP_1_2_2_5ha.tif,  PAD_GAP_1_2_5ha.tif to the FORESCEA1B 2050 eco4 level3 
```{r}
GAP12_6.25ha <- rast(here("PAD_1_2_rasters","PAD_GAP_1_2_6.25ha.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(GAP12_6.25ha) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_6.25ha <- FORESCE_A1B %>%
  crop(y = GAP12_6.25ha) %>%
  mask(mask = GAP12_6.25ha)

plot(foresce_pad_6.25ha)

# save raster as .tif
terra::writeRaster(
  foresce_pad_6.25ha,
  "FORESCE_2050_A1B_us_extent_6.25ha.tif",
  overwrite = TRUE
  )


## now we now we can crop that new FORESCE_2050_A1B_us_extent_[ha].tif to the FORESCE2050A1B_eco_4_Level_3_2013.tif

FORESCE_A1B_eco4lev3_us_6.25ha <- rast(here("FORESCE_PAD_6.25ha", "FORESCE_2050_A1B_us_extent_6.25ha.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_A1B_eco4lev3_us_6.25ha) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_A1B_eco4lev3_us_6.25ha %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3_6.25ha.tif",
  overwrite = TRUE
  )


```




## difference?
```{r}
newtest = raster_5ha - raster_1ha
plot(newtest)

```


```{r}
PAD_GAP_1_2_US_5ha <- sf::st_read(here("PAD","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US.shp")) %>% 
  dplyr::filter(SHAPE_Area > 50000) %>% # filter for geometries greater than 1ha
  dplyr::select(GAP_Sts)

# plot(PAD_GAP_1_2_US_1ha)

template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_GAP_1_2_US_5ha <- st_transform(
  PAD_GAP_1_2_US_5ha,
  crs=st_crs(template)
)


PAD_GAP_1_2_US_5ha$GAP_Sts = as.numeric(PAD_GAP_1_2_US_5ha$GAP_Sts)

raster <- fasterize(
  sf = PAD_GAP_1_2_US_5ha,
  raster = template,
field = "GAP_Sts")

plot(raster)

raster::writeRaster(
   raster,
   "PAD_GAP_1_2_5ha_v2.tif",
   overwrite = TRUE
   )

## FORESCE LULC
GAP12_5ha <- rast(here("PAD_1_2_rasters","PAD_GAP_1_2_5ha_v2.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(GAP12_5ha) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_5ha <- FORESCE_A1B %>%
  crop(y = GAP12_5ha) %>%
  mask(mask = GAP12_5ha)

# plot(foresce_pad_6.25ha)

# save raster as .tif
terra::writeRaster(
  foresce_pad_5ha,
  "FORESCE_2050_A1B_us_extent_5ha_v2.tif",
  overwrite = TRUE
  )

## now crop to SE US
FORESCE_A1B_eco4lev3_us_1ha <- rast(here("FORESCE_PAD_1ha", "FORESCE_2050_A1B_us_extent_1ha_v2.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_A1B_eco4lev3_us_1ha) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_A1B_eco4lev3_us_1ha %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3_1ha_v2.tif",
  overwrite = TRUE
  )

```



```{r}
PAD_GAP_1_2_US<- sf::st_read(here("PAD","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US.shp")) 

# reg4_eco_l4 <- sf::st_read(here("reg4_eco_l4","reg4_eco_l4.shp")) %>% 
#  st_transform(st_crs(PAD_GAP_1_2_US))


PAD_GAP_1_2_US_dissolve_clip <- sf::st_read(here("PAD_dissolve_clip","PADUS2_1Combined_Proclamation_Fee_Designation_Easement_GAP1_2_US_dissolve_Unit_Nm_clip_eco4.shp")) 

# geom_histogram(PAD_GAP_1_2_US_dissolve$SHAPE_Area)

PAD_df<- data_frame(PAD_GAP_1_2_US_dissolve_clip$Unit_Nm)  %>% 
  mutate(Category = PAD_GAP_1_2_US_dissolve_clip$Category,
         Loc_Mang = PAD_GAP_1_2_US_dissolve_clip$Loc_Mang,
         State_Nm = PAD_GAP_1_2_US_dissolve_clip$State_Nm,
         GIS_Acres = PAD_GAP_1_2_US_dissolve_clip$GIS_Acres,
         GAP_Sts = PAD_GAP_1_2_US_dissolve_clip$GAP_Sts,
    Unit_Nm = PAD_GAP_1_2_US_dissolve_clip$Unit_Nm,
         shape_leng = PAD_GAP_1_2_US_dissolve_clip$SHAPE_Leng,
         shape_area = PAD_GAP_1_2_US_dissolve_clip$SHAPE_Area,
         geometry = PAD_GAP_1_2_US_dissolve_clip$geometry) 
 
options(scipen=999)

hist(PAD_df$shape_leng, breaks = 20, col = "lightblue",
     ylim=c(0,3500),
     xaxp = c(0, 1200000, 10))

hist(PAD_df$shape_area, breaks = 20, col = "lightblue",
     ylim=c(0,3500),
     xaxp = c(0, 400000000, 10))

hist(PAD_df$GIS_Acres, breaks = 20, col = "lightblue",
     ylim=c(0,3500))

```

```{r}
PAD_df_filt <- PAD_df %>% 
  select(-`PAD_GAP_1_2_US_dissolve_clip$Unit_Nm`) %>% 
  filter(shape_area >= 500000) %>% 
  arrange(desc(shape_area)) %>% 
  slice_max(shape_area, n = 927) ## selecting top 50% of shape area geometrys 

hist(PAD_df_filt$shape_area, breaks = 20, col = "lightblue")

sf::st_write(PAD_df_filt,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/PAD_df_filt/PAD_df_filt.shp") # export df as shp to view in qgis 

```

## check to see if GIS_Area column is comparable
## yes 1855 vs 1844 observarions >50 ha or 123.44 acres 
```{r}
# PAD_df_filt_2 <- PAD_df %>% 
#   select(-`PAD_GAP_1_2_US_dissolve_clip$Unit_Nm`) %>% 
#   filter(GIS_Acres >= 123.55) %>% 
#   arrange(desc(GIS_Acres)) ## selecting top 50% of shape area geometrys 
```


## now lets rasterize the shp file with the FORESCE tif as template
```{r}
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_filt_shp <- sf::st_read(here("PAD_df_filt","PAD_df_filt.shp"))

PAD_df_filt_shp <- st_transform(
  PAD_df_filt_shp,
  crs=st_crs(template)
)

PAD_df_filt_shp$GAP_Sts = as.numeric(PAD_df_filt_shp$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_filt_shp_shp,
  raster = template,
field = "GAP_Sts")

plot(raster)

raster::writeRaster(
   raster,
   "PAD_filt.tif",
   overwrite = TRUE
   )

## FORESCE LULC
PAD_filt <- rast(here("PAD_filt.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt) %>%
  mask(mask = PAD_filt)

# plot(foresce_pad_filt)

# save raster as .tif
terra::writeRaster(
  foresce_pad_filt,
  "FORESCE_2050_A1B_us_extent_50ha_filt.tif",
  overwrite = TRUE
  )

## now crop to SE US
FORESCE_2050_A1B_us_extent_50ha <- rast(here("FORESCE_2050_A1B_us_extent_50ha_filt.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_50ha) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_50ha %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3_50ha_filt.tif",
  overwrite = TRUE
  )
```

## lets do the same thing now but for 25ha 
```{r}
PAD_df_filt_25ha <- PAD_df %>% 
  select(-`PAD_GAP_1_2_US_dissolve_clip$Unit_Nm`) %>% 
  filter(shape_area >= 250000) %>% 
  arrange(desc(shape_area)) %>% 
  slice_max(shape_area, n = 1145) ## selecting top 50% of shape area geometrys 


hist(PAD_df_filt_25ha$shape_area, breaks = 20, col = "lightblue")

sf::st_write(PAD_df_filt_25ha,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/PAD_df_filt_25ha/PAD_df_filt_25ha.shp") # export df as shp to view in qgis 

```

## now lets rasterize the 25ha PAD shp file with the FORESCE tif as template and then crop and mask
```{r}
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_filt_25ha_shp <- sf::st_read(here("PAD_df_filt_25ha","PAD_df_filt_25ha.shp"))

PAD_df_filt_25ha_shp <- st_transform(
  PAD_df_filt_25ha_shp,
  crs=st_crs(template)
)

PAD_df_filt_25ha_shp$GAP_Sts = as.numeric(PAD_df_filt_25ha_shp$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_filt_25ha_shp,
  raster = template,
field = "GAP_Sts")

plot(raster)

raster::writeRaster(
   raster,
   "PAD_filt_25ha.tif",
   overwrite = TRUE
   )

## FORESCE LULC
PAD_filt_25ha <- rast(here("PAD_filt_25ha.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt_25ha) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt_25ha) %>%
  mask(mask = PAD_filt_25ha)

# plot(foresce_pad_filt)

# save raster as .tif
terra::writeRaster(
  foresce_pad_filt,
  "FORESCE_2050_A1B_us_extent_25ha_filt.tif",
  overwrite = TRUE
  )

## now crop to SE US
FORESCE_2050_A1B_us_extent_25ha <- rast(here("FORESCE_2050_A1B_us_extent_25ha_filt.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_25ha) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_25ha %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3_25ha_filt.tif",
  overwrite = TRUE
  )
```

## filter for bobcat range, 23km^2 based on https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0194243 
## other papers: 95.7 km2  based on https://www.jstor.org/stable/3801498?casa_token=KXmnO1z2bAIAAAAA%3A1kcBM2luAbYn3fyNBOutIAuYG8K6Md9PIJJSs5JrNoJkXmSwYPHlskgyMl34QAO4O_GXNBz623KTnWZo7Jzoq3UcS9CxUYb8R8n09JKy7N_7Z9Y8klSm&seq=1
## another paper: 57.3 km2  based on https://www.sciencedirect.com/science/article/abs/pii/S0006320711002539 
```{r}
PAD_df_filt <- PAD_df %>% 
  select(-`PAD_GAP_1_2_US_dissolve_clip$Unit_Nm`) %>% 
  filter(shape_area >=63450000) %>% 
  arrange(desc(shape_area))  ## n = 230 

# hist(PAD_df_filt$shape_area, breaks = 20, col = "lightblue")

sf::st_write(PAD_df_filt,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/PAD_df_filt/PAD_df_BCrange_63km.shp") # export df as shp to view in qgis 

```


## now lets rasterize the bobcat home range shps and include the bobcat home range PAD tif with the FORESCE tif as template and then crop and mask
```{r}
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_filt_63km_shp <- sf::st_read(here("PAD_df_filt","PAD_df_BCrange_63km.shp"))

PAD_df_filt_63km_shp <- st_transform(
  PAD_df_filt_63km_shp,
  crs=st_crs(template)
)

PAD_df_filt_63km_shp$GAP_Sts = as.numeric(PAD_df_filt_63km_shp$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_filt_63km_shp,
  raster = template,
field = "GAP_Sts")

plot(raster)

raster::writeRaster(
   raster,
   "PAD_filt_63km.tif",
   overwrite = TRUE
   )


## FORESCE LULC
PAD_filt_63km <- rast(here("PAD_filt_63km.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt_63km) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt_63km) %>%
  mask(mask = PAD_filt_63km)

# plot(foresce_pad_filt)

# save raster as .tif
terra::writeRaster(
  foresce_pad_filt,
  "FORESCE_2050_A1B_us_extent_63km_filt.tif",
  overwrite = TRUE
  )

## now crop to SE US
FORESCE_2050_A1B_us_extent_63km <- rast(here("FORESCE_2050_A1B_us_extent_63km_filt.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_63km) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_63km %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

# plot(crop)

# save raster as .tif
terra::writeRaster(
  crop,
  "FORESCE_2050_A1B_eco4_l3_63km_filt.tif",
  overwrite = TRUE
  )
```
## filter bobcat range by Litvaitis et al., 1986 (maine study)  male bobcats range 95.7km2 (9.57e+7 m2)
```{r}
PAD_df_filt_96km <- PAD_df %>% 
  select(-`PAD_GAP_1_2_US_dissolve_clip$Unit_Nm`) %>% 
  filter(shape_area >=9.57e+7) %>% 
  arrange(desc(shape_area))  ## n = 44

# hist(PAD_df_filt$shape_area, breaks = 20, col = "lightblue")

sf::st_write(PAD_df_filt_96km,"~/Desktop/FORESCE_PAD/FORESSCE_PAD/PAD_df_filt/PAD_df_BCrange_96km.shp") # export df as shp to view in qgis 
```

## now lets rasterize the bobcat home range shps and include the bobcat home range PAD tif with the FORESCE tif as template and then crop and mask
```{r}
template <- raster(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif")) 

PAD_df_filt_96km_shp <- sf::st_read(here("PAD_df_filt","PAD_df_BCrange_96km.shp"))

PAD_df_filt_96km_shp <- st_transform(
  PAD_df_filt_96km_shp,
  crs=st_crs(template)
)

PAD_df_filt_96km_shp$GAP_Sts = as.numeric(PAD_df_filt_96km_shp$GAP_Sts)

raster <- fasterize(
  sf = PAD_df_filt_96km_shp,
  raster = template,
field = "GAP_Sts")

plot(raster)

# raster::writeRaster(
#    raster,
#    "PAD_filt_96km.tif",
#    overwrite = TRUE
#    )


## FORESCE LULC
PAD_filt_96km <- rast(here("PAD_filt_96km.tif"))
FORESCE_A1B <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013_us_extent.tif"))

# set crs so they're the same 
crs(PAD_filt_96km) <-  "EPSG:5070"
crs(FORESCE_A1B) <-  "EPSG:5070"

foresce_pad_filt <- FORESCE_A1B %>%
  crop(y = PAD_filt_96km) %>%
  mask(mask = PAD_filt_96km)

# plot(foresce_pad_filt)

# save raster as .tif
# terra::writeRaster(
#   foresce_pad_filt,
#   "FORESCE_2050_A1B_us_extent_96km_filt.tif",
#   overwrite = TRUE
#   )

## now crop to SE US
FORESCE_2050_A1B_us_extent_96km <- rast(here("FORESCE_2050_A1B_us_extent_96km_filt.tif"))
FORESCE_A1B_eco4lev3 <- rast(here("FORESCE2050A1B_eco_4_Level_3_2013.tif"))

crs(FORESCE_2050_A1B_us_extent_96km) <-  "EPSG:5070"
crs(FORESCE_A1B_eco4lev3) <-  "EPSG:5070"

crop <- FORESCE_2050_A1B_us_extent_96km %>%
  crop(y = FORESCE_A1B_eco4lev3) %>%
  mask(mask = FORESCE_A1B_eco4lev3)

# plot(crop)

# save raster as .tif
# terra::writeRaster(
#   crop,
#   "FORESCE_2050_A1B_eco4_l3_96km_filt.tif",
#   overwrite = TRUE
#   )
```
